The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Revisit DAS caching system subsection. Added description of the algorithm
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@646 4525493e-7705-40b1-a816-d608a930855b

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS caching system}
The DAS caching system was used to store data-service results and aggregated
information in form of DAS records. As it shown in figure \ref{DAS_cache} 
the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
and aggregate data over there. The independent set of daemons, DAS robots, were
constantly monitor Analytics DB for most popular requests and pre-fetch most
common data into DAS cache.

Each user request placed to the system was parsed and set of selection
and condition keys were identified. They were mapped into set of
data service APIs and their input parameters. DAS
cache was checked for a queries with a similar set of selection and condition
keys used earlier.\footnote{Since DAS didn't hold
any data by default we cannot rely on data being present in a cache.
Moreover if we look-up data first, we cannot tell if it was complete
set or not, since earlier queries may retrieve only portion of the requested data.}
If the superset was found a data were
retrieved from the cache, otherwise a data-service APIs were
invoked and the results were collected and aggregated inside of DAS cache.

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys known to DAS to describe common
entities of our data. For example, {\it dataset, file, block, run, event}. Those
names were used by our users in their daily work operations. So, by typing
{\it file} we didn't perform a string matching to ``file'' string, but rather
look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
For intance, for {\it dataset, block, file} the default attribute 
was a {\it name}, while for {\it run} and {\it event} the default attribute was 
a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
{\it run=123456}, which means run number 123456. At the same time a queyr 
key, e.g. {\it site} meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as {\it run=123456} instruct system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

@@

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system and data-service;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
Figure \ref{db_size} shows the current size of meta-data from aforementioned 
data-services.
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{db_size.pdf}
\caption{
Current size of meta-data for CMS data-services participated in DAS, before
official data-taking. 
%We anticipate to collect $\sim$500GB of meta-data
%per year during normal LHC operations.
}
\label{db_size}
\end{figure}

>>>>

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=120mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{description}
\item[  1.] Decompose query into set of conditions ($conds$) 
and selection keys ($skeys$)
\item[  2.] Check cache for query hash, if found: look-up results from cache
\item[  3.] Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
\item[  4.] Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics_found(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
\item[  5.] Perform aggrgation by iterating over the previos step
\begin{verbatim}
if cache_has(row.match_key):
   existing_record = get_from_cache(row.match_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
\item[  6.] Get results from the cache
\end{description}
The data was stored in a form of DAS records into the cache, while
queries were recorded into Analytics DB for further analysis. We
also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

%The DAS caching system was used to store data-service results and aggregated
%information in form of DAS records. As it shown in figure \ref{DAS_cache} 
%the DAS cache server updates Analytics DB, requests and retrieve data from DAS cache
%and aggregate data over there. The independent set of daemons, DAS robots, were
%constantly monitor Analytics DB for most popular requests and pre-fetch most
%common data into DAS cache.

%Each user request placed to the system was parsed and set of selection
%and condition keys were identified. They were mapped into set of
%data service APIs and their input parameters. DAS
%cache was checked for a queries with a similar set of selection and condition
%keys used earlier.\footnote{Since DAS didn't hold
%any data by default we cannot rely on data being present in a cache.
%Moreover if we look-up data first, we cannot tell if it was complete
%set or not, since earlier queries may retrieve only portion of the requested data.}
%If the superset was found a data were
%retrieved from the cache, otherwise a data-service APIs were
%invoked and the results were collected and aggregated inside of DAS cache.

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys known to DAS to describe common
entities of our data. For example, {\it dataset, file, block, run, event}. Those
names were used by our users in their daily work operations. So, by typing
{\it file} we didn't perform a string matching to ``file'' string, but rather
look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
For intance, for {\it dataset, block, file} the default attribute 
was a {\it name}, while for {\it run} and {\it event} the default attribute was 
a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
{\it run=123456}, which means run number 123456. At the same time a query 
key, e.g. {\it site} meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as {\it run=123456} instruct system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

@@

\section{Results\label{Results}}
In this section we discuss preliminary restults achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS file transfer system and data-service;
\item SiteDB \cite{SiteDB}, the CMS site resources data-service;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiar properties. 
For instance, the data in PhEDEx are transient due to constant 
migration of CMS data; the DBS system were divided into dozen 
of individual instances with own size and lifetime, who has been used by different production 
teams for various bookkeeping tasks; the RunSummary data-service
operates on private network accessible at CMS online operation center.
Table \ref{db_size} shows the current size of meta-data from aforementioned 
data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
\hline
tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
\hline
\hline
\end{tabular}
\caption{
Current size of meta-data for CMS data-services participated in DAS, before
official data-taking. 
}
\label{db_size}
\end{table*}
